% abstract.tex
\placeholder{Rewrite in finished prose once Phase~2 (theorem proofs)
lands.  v0.2 outline: states the diamond-progression construction, the
three result-density theorems, the dimensional-selection corollary,
and the deferred-to-motivating-examples scope of Boltzmann + NS.}

\noindent
Paper~I of the A=1 Discrete Causal Lattice (DCL) series established
the bipartite octahedral lattice as the substrate of fundamental
physics in (3+1) dimensions.  This paper -- the foundational
mathematics companion -- generalises that lattice to all dimensions
$d \geq 2$ via a single construction (the \emph{diamond progression}:
the d-of-(d+1) regular simplex vertex set), proves three result-
density theorems characterising the resulting geometry, and derives a
corollary that selects $d = 3$ as the framework's admissible
dimension.

\medskip

\noindent
The three theorems are
\begin{itemize}
\item $\text{coord}(d) = 2d$ (linear in dimension),
\item maximum Euclidean hop = $\sqrt{d}$,
\item unit-cell volume = $(d+1)^{(d-1)/2}$.
\end{itemize}
The linear-coordination result distinguishes the diamond progression
from chemical-diamond / face-centred-cubic-style constructions where
coordination stays bounded; the consequence is that
propagation through the $d=3$ lattice is \emph{denser} (more hops
per unit Euclidean distance) than through higher-$d$ instances.

\medskip

\noindent
The headline corollary couples the diamond progression to a
hypothetical minimum probability quantum $\delta p_\min$:

\begin{equation*}
d_{\max} = \frac{1}{\delta p_\min} - 1.
\end{equation*}
At $\delta p_\min = 1/4$, $d_{\max} = 3$ -- i.e.\ the lattice
\emph{cannot} sustain the A=1 unity axiom in $d \geq 4$ because
some node would have to carry less than $\delta p_\min$.  This is the
candidate mechanism that \emph{forces} the universe to be three
dimensional rather than permitting it.  The empirical investigation of
$\delta p_\min$ -- including the question of whether the framework
forces the natural value $1/4$ -- lives in the companion repository
\texttt{dcl-delta-p-min}.

\medskip

\noindent
The remainder of the paper develops the analytic foundations
(discrete differential geometry, function spaces, discrete measure
theory) on the diamond progression and constructs the discrete-to-
continuum limit machinery (Section~\ref{sec:continuum_limits})
that the planned Hilbert-Sixth capstone consumes.  Boltzmann-shaped
operators and the lattice-Boltzmann $\to$ NS / Euler fluid limit
appear as \emph{motivating examples}
(Section~\ref{sec:motivating_examples}); their full proofs are
deferred to subsequent papers.

\medskip

\noindent
The paper's status across its claims is tracked in
Section~\ref{sec:introduction}, Table~\ref{tab:audit}.  At
v0.2-DRAFT, the inherited Paper~I guardrail rows are \texttt{PASS};
the three result-density theorems and the dimensional-selection
corollary are \texttt{STUB} pending Phase~2 proofs.
