% introduction.tex
% Foundations-of-Paper-IV introduction (v0.2 recalibration).  Foregrounds
% the diamond-progression construction, the three result-density theorems,
% and the dimensional-selection corollary as the paper's spine.

\label{sec:introduction}

\placeholder{Rewrite in finished prose once Phase~2 sections land.
At v0.2 the introduction states the problem (extending Paper~I's
bipartite lattice to all $d$), the construction (diamond progression),
the three theorems, and the dimensional-selection corollary that
makes the paper structurally complete.}

\subsection{Problem}
\label{subsec:problem}

Paper~I~\cite{example2024} established that the bipartite octahedral
lattice carries the (3+1)-D Lorentz structure of fundamental
physics.\footnote{Citations through this draft refer to placeholder
BibTeX entries in \texttt{paper-bib/references.bib}; real cites land
as the prose is written.}  The construction in Paper~I is specific
to $d = 3$ -- the physical setting -- but the mathematics is
naturally dimension-agnostic.  Downstream work in the DCL series
needs the lattice at arbitrary $d$:

\begin{itemize}
\item Paper~II's gauge-derivation programme uses lattice symmetries
      whose statements do not depend on $d = 3$;
\item the planned Hilbert-Sixth axiomatisation capstone needs
      $d$-dimensional theorems for its universality claim;
\item future kinetic-theory work on the lattice (Boltzmann / Navier-
      Stokes / Euler) needs discrete-to-continuum limit machinery
      whose natural statement is dimension-agnostic.
\end{itemize}

But the more interesting question is not just \emph{can we extend the
mathematics to all $d$?} (yes; the construction below works) but
\emph{does the framework itself permit, prefer, or force a specific
$d$?}  We answer the second question affirmatively: combined with a
minimum probability quantum $\delta p_\min$, the framework forces
$d \leq 1/\delta p_\min - 1$.

\subsection{Contribution}
\label{subsec:contribution}

Four contributions, in order:

\paragraph{1.~The diamond progression.}
Section~\ref{sec:lattice_n_dim} introduces a single combinatorial
construction -- the d-of-(d+1) regular simplex vertex set -- that
generates a bipartite octahedral lattice in every dimension
$d \geq 2$.  The bipartition is strict, indexed by parity (even /
odd integers).  At $d = 3$ the construction recovers Paper~I's
substrate lattice exactly.  We call this family the \emph{diamond
progression} -- a custom variant, distinct from chemical diamond.

\paragraph{2.~Three result-density theorems.}
Section~\ref{sec:three_theorems} proves three theorems
characterising the geometry of the diamond progression as a
function of $d$:

\begin{itemize}
\item $\text{coord}(d) = 2d$,
\item maximum Euclidean hop = $\sqrt{d}$,
\item unit-cell volume = $(d+1)^{(d-1)/2}$.
\end{itemize}
The first is the most consequential: it distinguishes the diamond
progression from chemical-diamond-style constructions (which stay at
coord~$4$ as $d$ grows) and underwrites the dimensional-selection
corollary in Section~\ref{sec:dimensional_selection}.

\paragraph{3.~The Hilbert-Sixth on-ramp.}
Sections~\ref{sec:discrete_diff_geometry}--\ref{sec:continuum_limits}
develop the analytic foundations on the diamond progression:
discrete differential geometry, function spaces, discrete measure
theory, and the discrete-to-continuum limit operators.  Section~\ref{sec:continuum_limits}
is the load-bearing section the planned Hilbert-Sixth capstone
consumes.

\paragraph{4.~Dimensional selection.}
Section~\ref{sec:dimensional_selection} couples the diamond
progression to a hypothetical $\delta p_\min$ and derives
$d_{\max} = 1/\delta p_\min - 1$.  At $\delta p_\min = 1/4$,
$d_{\max} = 3$.  The result is a candidate mechanism that
\emph{forces} the universe to be three dimensional.  The empirical
investigation of $\delta p_\min$ (and the question of whether the
framework forces the value $1/4$) lives in the companion repository
\texttt{dcl-delta-p-min}; this paper provides the mathematical
statement on the assumption that some $\delta p_\min > 0$ exists.

\subsection{Scope}
\label{subsec:scope}

Out of scope for this first paper:

\begin{itemize}
\item \emph{Physics applications.}  The lattice's role as substrate
      of fundamental physics is the subject of Papers~I, II, III and
      the planned discrete-probability paper.
\item \emph{Sympy operationalisation.}  Machine-checked versions of
      the three theorems and the dimensional-selection corollary
      live in the planned \texttt{dcl-formalism} package.
\item \emph{Boltzmann / NS proof.}  Sections~\ref{sec:motivating_examples}
      states the lattice kinetic theory and fluid-limit ansatz as
      motivating examples; full proofs are deferred to subsequent
      papers.
\item \emph{Pinning $\delta p_\min$'s value.}  Whether
      $\delta p_\min = 1/4$ is forced by the framework, by
      observational constraints, or by both is investigated in
      \texttt{dcl-delta-p-min} (Phase~1 theoretical pass) and
      plausibly in the planned discrete-probability paper.  This
      paper takes $\delta p_\min > 0$ as a hypothesis and derives
      the $d_{\max}$ corollary; the value is symbolic.
\item \emph{The aspirational sequel `A=1 forces linear coordination'}
      (working name F4 in the user's `raw_thoughts.md`): the claim
      that, among bipartite lattices, only the diamond-progression
      family ($\text{coord} = 2d$) admits A=1 non-degenerately.  A
      conjecture for a follow-up paper, not a result here.
\item \emph{The aspirational generator-fractal unification}
      (`prize-of-prizes' in `raw_thoughts.md`): a single
      IFS / Bratteli diagram / operad generating the diamond
      progression as a fractal across $d$.  Aspirational; future
      work.
\end{itemize}

The audit table (Table~\ref{tab:audit}) is the canonical answer to
``has this claim been proved yet?''.

%% The audit table is included here so the reader meets it before any
%% claim that depends on a status entry.
\input{sections/audit_table}
