% motivating_examples.tex
% Section 10 -- motivating examples (framing only).  Boltzmann-shaped
% collision operators on the lattice and the lattice-Boltzmann ->
% NS / Euler fluid-limit ansatz.  Full proofs are out of scope for
% this paper; deferred to subsequent papers in this repo or in sibling
% repos.

\label{sec:motivating_examples}

\placeholder{Section body -- write during Phase 5.  This section is
\emph{framing only}: it states what the foundations toolkit
developed in Sections~\ref{sec:lattice_n_dim}--\ref{sec:continuum_limits}
will eventually be used for.  Full proofs of the H-theorem on
$\Lambda_d$ and of the fluid-limit convergence are deferred to
subsequent papers.}

\subsection{Boltzmann-Shaped Operators on the Diamond Progression}
\label{subsec:boltzmann}

\placeholder{Define a Boltzmann-shaped collision operator on
$\Lambda_d$ (or its rescaling $\epsilon \Lambda_d$): a binary /
$k$-ary collision rule that respects the bipartite structure and
the automorphism algebra of
Section~\ref{sec:automorphism_algebra}.  Define the lattice
$H$-functional.  State (without proving) the expected $H$-theorem
and the expected equilibrium measures.  Cross-reference to a future
paper for the full proof.}

\paragraph{Motivating-example status.}
\todomark{motivating-example: the $H$-theorem on $\Lambda_d$ is
asserted here but not proved.  Full proof scheduled for a
subsequent paper.  See Table~\ref{tab:audit}, ``Boltzmann-shaped
operators'' row, which is by design \texttt{STUB}.}

\subsection{Fluid (Navier-Stokes / Euler) Limit Ansatz}
\label{subsec:fluid_limit}

\placeholder{State the lattice-Boltzmann $\to$ NS / Euler limit
ansatz: the scaling of $\epsilon$ and of the collision-rule
parameters that should recover the continuum fluid equations in
the limit (via the convergence machinery of
Section~\ref{sec:continuum_limits}).  Identify the technical
ingredients the convergence proof would need.  Sketch the
expected statement.  Cross-reference to a future paper for the
full proof.}

\paragraph{Motivating-example status.}
\todomark{motivating-example: the lattice-Boltzmann $\to$ NS / Euler
convergence is asserted here but not proved.  Full proof scheduled
for a subsequent paper.  See Table~\ref{tab:audit}, ``Fluid
(NS / Euler) limit ansatz'' row, which is by design \texttt{STUB}.}

\subsection{Why These Examples}
\label{subsec:why_these_examples}

\placeholder{Hilbert's 1900 problem singled out Boltzmann's kinetic
theory and the rigorous Boltzmann $\to$ NS / Euler limit as the
centerpiece of an axiomatised physics.  Treating them as motivating
examples here -- rather than as proof targets of this paper --
keeps this first paper focused on the foundations toolkit while
making the eventual use of that toolkit visible.  The full proofs
of the $H$-theorem on $\Lambda_d$ and of the fluid-limit
convergence are scheduled for subsequent papers in this repository
or in sibling repositories.}
