\documentclass[12pt, letterpaper]{article}

%% Build from the paper/ directory:
%%   cd paper && pdflatex main.tex
%% or use the root build system:
%%   make paper          (from repo root)
%%   ./build.sh paper    (POSIX)
%%   build.cmd paper     (Windows cmd)

\input{macros/packages}
\input{macros/commands}

%% Page geometry (also set here as a fallback in case the package path fails)
\geometry{top=1in, bottom=1in, left=1in, right=1in}

%% ---------------------------------------------------------------
%% TITLE PAGE
%% ---------------------------------------------------------------

\title{Foundations of the A=1 Discrete Causal Lattice:\\
       the Diamond Progression in Arbitrary Dimension%
       \thanks{Version 0.2-DRAFT.
       \href{https://github.com/JackDMenendez/dcl-mathematics}%
            {GitHub: JackDMenendez/dcl-mathematics}.
       %% Fill in at release time (conditional on publication decision):
       %% \href{https://doi.org/10.5281/zenodo.XXXXXXXX}%
       %%      {DOI: 10.5281/zenodo.XXXXXXXX}.
       Pre-release draft; DOI to be assigned if and when an external
       publication decision is taken.}}
\author{Jack Menendez\\
        \small Independent Scholar, Eugene, OR\\
        \small ORCID: \href{https://orcid.org/0009-0003-1166-307X}{0009-0003-1166-307X}\\
        \small \href{mailto:jackdmenendez@gmail.com}{jackdmenendez@gmail.com}}
\date{\today}

\begin{document}

\maketitle

%% ---------------------------------------------------------------
%% FRONT-MATTER NOTICES
%% ---------------------------------------------------------------
\noindent\textbf{License.}  Paper text and figures: Creative Commons
Attribution 4.0 (CC BY 4.0).  Source code in the companion
repository: MIT License.

\medskip
\noindent\textbf{Data availability.}  The LaTeX source for this paper
is publicly available at
\href{https://github.com/JackDMenendez/dcl-mathematics}%
     {\nolinkurl{github.com/JackDMenendez/dcl-mathematics}}.
This is a paper-only repository in the A=1 Discrete Causal Lattice
(DCL) series; no numerical experiments accompany the first paper.
%% At release time (if external publication proceeds):
%% The exact commit corresponding to this release is the vX.Y tag,
%% archived at
%% \href{https://doi.org/10.5281/zenodo.XXXXXXXX}%
%%      {\nolinkurl{doi:10.5281/zenodo.XXXXXXXX}}.

\medskip
\noindent\textbf{Series position.}  Supporting-mathematics repository
of the A=1 DCL series.  Upstream of the planned
\texttt{dcl-formalism} sympy package (which operationalises selected
theorems from this paper as machine-checked primitives) and of the
planned Hilbert-Sixth capstone paper (Paper~I follow-on \#18, which
consumes the discrete-to-continuum limit machinery of
Section~\ref{sec:continuum_limits} for its axiomatisation claim).
Downstream of Paper~I (\emph{Geometry First}, doi:10.5281/zenodo.20078529),
the source-of-record for the bipartite octahedral lattice in its
physical (3+1)-D setting.

\medskip
\noindent\textbf{Status legend (used throughout, including the audit
table in Section~\ref{sec:introduction}).}  \texttt{PASS} -- derivation
complete or experiment confirms the claim to stated precision.
\texttt{PART} -- mechanism demonstrated, full quantitative match
pending; specific gaps documented in the evidence column.
\texttt{STUB} -- analytical result with no full proof yet, or proof
planned but not yet written.  \texttt{FAIL} -- tested and
disconfirmed.

\medskip
\noindent\textbf{Keywords.}  discrete causal lattice; bipartite
octahedral lattice; discrete differential geometry; discrete-to-
continuum limit; Hilbert's sixth problem; foundations of physics.

\newpage

\begin{abstract}
    \input{sections/abstract}
\end{abstract}

\newpage
\tableofcontents
\newpage

%% ============================================================
%% PART I: COMBINATORIAL FOUNDATIONS (THE DIAMOND PROGRESSION)
%% ============================================================
\section{Introduction}
\input{sections/introduction}

\section{The Diamond Progression: d-of-(d+1) Simplex Vertex Construction}
\input{sections/lattice_n_dim}

\section{Three Result-Density Theorems}
\input{sections/three_theorems}

\section{Automorphism Algebra in Dimension $d$}
\input{sections/automorphism_algebra}

%% ============================================================
%% PART II: ANALYTIC FOUNDATIONS
%% ============================================================
\section{Discrete Differential Geometry on the Diamond Progression}
\input{sections/discrete_diff_geometry}

\section{Function Spaces on the Diamond Progression}
\input{sections/function_spaces}

\section{Discrete Measure Theory on the Diamond Progression}
\input{sections/discrete_measure_theory}

%% ============================================================
%% PART III: THE HILBERT-SIXTH ON-RAMP
%% ============================================================
\section{Discrete-to-Continuum Limit Operators}
\input{sections/continuum_limits}

%% ============================================================
%% PART IV: DIMENSIONAL SELECTION (HEADLINE COROLLARY)
%% ============================================================
\section{Dimensional Selection: $d_{\max} = 1/\delta p_{\min} - 1$}
\input{sections/dimensional_selection}

%% ============================================================
%% PART V: MOTIVATING EXAMPLES (FRAMING ONLY)
%% ============================================================
\section{Motivating Examples: Boltzmann and the Fluid Limit}
\input{sections/motivating_examples}

\section{Conclusion}
\input{sections/conclusion}

\section*{Acknowledgements}
\input{sections/acknowledgements}

%% ============================================================
%% APPENDICES
%% ============================================================
\appendix

\section{Symbols}
\input{sections/symbols}

\section{Terms}
\input{sections/terms}

\section{Code and Data}
\input{sections/code_and_data}

\section{Reproducibility Quickstart}
\input{sections/reproducibility}

%% ============================================================
\bibliographystyle{unsrt}
\bibliography{paper-bib/references}

\end{document}
